---
title: "Ops let's write title later"
author: "Michael Zhou, Nigel Li, Stella Pham"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(dplyr)
library(lubridate)
library(knitr)
library(ggplot2)
```

```{r eval = TRUE, echo = FALSE, message = FALSE}
raw_data = read.csv("data/fueltypesall.csv")
```

## Section 1: Introduction
### Background of the Data
This data set contains weekly retail prices for six fuel types: Regular Unleaded 
Gasoline, Mid‑Grade Gasoline, Premium Gasoline, Diesel, Auto Propane, and 
Compressed Natural Gas across 10 markets in Ontario. Prices are expressed in 
Canadian cents per litre, with compressed natural gas measured in cents per 
gasoline-equivalent litre. The data set spans from January 3, 1990 to March 16, 
2026 and is updated weekly with the latest retail fuel price data published by 
the Government of Ontario.

### Research Question
How does the price differ between cities in Ontario?

## Section 2: Data Description
The data set contains 11,252 weeks of data from 1990 to March 16, 2026, covering 
six fuel types. Each row represents a weekly average of fuel prices, so daily 
fluctuations are not captured. It also includes three regional average columns 
(Ontario, Northern Ontario, Southern Ontario), which do not correspond to the 
simple average of the included cities, indicating they were sourced externally.

## Section 3: Data Preparation/Cleaning
Prior to analysis, it is important to highlight certain features of the raw 
data set that affect how it should be handled:

### 1. Bilingual Structure of the Data Set
The data set contains both French and English labels, sometimes separated by 
slashes or presented in different columns. To maintain consistency, 
variables are standardized into a single language format before analysis.

We first rename the columns we keep.

```{r eval = TRUE, echo = FALSE, message = FALSE}
d = raw_data %>%
  rename(
    Toronto_West = Toronto.West.Ouest,
    Toronto_East = Toronto.East.Est,
    St_Catharines = St..Catharine.s,
    Sault_Saint_Marie = Sault.Saint.Marie,
    Thunder_Bay = Thunder.Bay,
    North_Bay = North.Bay,
    Parry_Sound = Parry.Sound,
    Ontario_Average = Ontario.Average.Moyenne.provinciale,
    Southern_Ontario_Average = Southern.Average.Moyenne.du.sud.de.l.Ontario,
    Northern_Ontario_Average = Northern.Average.Moyenne.du.nord.de.l.Ontario,
    Fuel_Type = Fuel.Type
  )
```

```{r eval = TRUE, echo = FALSE, message = TRUE}
old_cols = names(raw_data)
new_cols = names(d)

names_comparison = data.frame(
  Before = old_cols,
  After = new_cols,
  stringsAsFactors = FALSE 
)

library(knitr)
kable(names_comparison, caption = "Old Column Names vs. New Column Names")
```

```{r eval = TRUE, echo = FALSE, message = FALSE}
d = d %>% 
  select(!Type.de.carburant)
```

We also remove the column "Type.de.carburant" because it was the 
same as the information in the Fuel_Type column. More specifically, it contained 
the fuel type names in French.

### 2. Handling of Zero Values
Values recorded as 0 in the city-level columns are interpreted as missing data, 
since these 0s do not represent actual prices. To prevent these zeros from 
affecting computations, all 0 values are converted to NA prior to analysis.

```{r eval = TRUE, echo = FALSE, message = FALSE}
d = d %>%
  mutate(across(where(is.numeric), ~ na_if(., 0)))
```

### 3. Conversion of the Date Column
The Week column, originally recorded as a character string, was converted to a 
Date type to ensure proper chronological ordering and accurate calculations. 
This allows for correct filtering, aggregation, and plotting over time, and 
ensures that weekly comparisons between cities are reliable.

```{r eval = TRUE, echo = FALSE, message = FALSE}
d = d %>%
  mutate(Date = as.Date(Date))
```

### 4. Categorization of the Data Set
As mentioned above, the data set includes six fuel types. To enable clearer 
and more accurate comparisons between cities, the data will be analyzed 
separately for each fuel type.

```{r eval = TRUE, echo = FALSE, message = FALSE}
regular_gas_data = d %>%
  filter(Fuel_Type == "Regular Unleaded Gasoline") %>%
  select(!Fuel_Type)

diesel_data = d %>%
  filter(Fuel_Type == "Diesel") %>%
  select(!Fuel_Type)

natural_gas_data = d %>%
  filter(Fuel_Type == "Compressed Natural Gas") %>%
  select(!Fuel_Type)

mid_gas_data = d %>%
  filter(Fuel_Type == "Mid-Grade Gasoline") %>%
  select(!Fuel_Type) 

premium_gas_data = d %>%
  filter(Fuel_Type == "Premium Gasoline") %>%
  select(!Fuel_Type)

auto_propane_data = d %>%
  filter(Fuel_Type == "Auto Propane") %>%
  select(!Fuel_Type)

```

## Section 4: Exploratory Data Analysis
### Visualizations

We want

```{r eval = TRUE, echo = FALSE, warning = FALSE}
d %>% 
  ggplot(aes(x = Date, y = Ontario_Average, color = Fuel_Type)) + 
  geom_line() +
  labs(
    title = "Weekly Fuel Prices in Ontario by Fuel Type",
    x = "Week",
    y = "Average Fuel Price (cents/litre)",
    color = "Fuel Type")
```

## Section 5: Statistical Inference
To answer our research question on how prices differ across cities, we will 
compare the Northern Ontario Average price to the Southern Ontario Average 
price. To ensure our analysis provides a comprehensive view of the 
transportation sector, we will focus our statistical inference on four distinct 
fuel categories: Regular Unleaded Gasoline (representing the majority of 
consumer vehicles), Diesel (representing commercial transit), as well as Auto 
Propane (representing alternative fuels). 
Additionally, we have restricted our analysis to recent data (2021–2026) to 
ensure our findings reflect the current economic landscape and provide the most 
relevant insights. Because prices are recorded on the exact same weeks, we will 
analyze the paired differences (Northern Average - Southern Average).
```{r eval = TRUE, echo = FALSE, message = FALSE}
# REGULAR GASOLINE
regional_compare_gas = regular_gas_data %>%
  filter(Date >= as.Date("2021-01-01")) %>%
  select(Date, Northern_Ontario_Average, Southern_Ontario_Average) %>%
  drop_na() %>%
  mutate(Reg_Diff = Northern_Ontario_Average - Southern_Ontario_Average)

t_test_gas = t.test(regional_compare_gas$Northern_Ontario_Average, 
                    regional_compare_gas$Southern_Ontario_Average,
                    paired = TRUE)
mean_diff_gas = t_test_gas$estimate
ci_lower_gas = t_test_gas$conf.int[1]
ci_upper_gas = t_test_gas$conf.int[2]
p_val_gas = t_test_gas$p.value

# DIESEL
regional_compare_diesel = diesel_data %>%
  filter(Date >= as.Date("2021-01-01")) %>%
  select(Date, Northern_Ontario_Average, Southern_Ontario_Average) %>%
  drop_na() %>%
  mutate(Reg_Diff = Northern_Ontario_Average - Southern_Ontario_Average)

t_test_diesel = t.test(regional_compare_diesel$Northern_Ontario_Average,
                       regional_compare_diesel$Southern_Ontario_Average,
                       paired = TRUE)
mean_diff_diesel = t_test_diesel$estimate
ci_lower_diesel = t_test_diesel$conf.int[1]
ci_upper_diesel = t_test_diesel$conf.int[2]
p_val_diesel = t_test_diesel$p.value

# AUTO PROPANE
regional_compare_propane = auto_propane_data %>%
  filter(Date >= as.Date("2021-01-01")) %>%
  select(Date, Northern_Ontario_Average, Southern_Ontario_Average) %>%
  drop_na() %>%
  mutate(Reg_Diff = Northern_Ontario_Average - Southern_Ontario_Average)

t_test_propane = t.test(regional_compare_propane$Northern_Ontario_Average,
                        regional_compare_propane$Southern_Ontario_Average,
                        paired = TRUE)
mean_diff_propane = t_test_propane$estimate
ci_lower_propane = t_test_propane$conf.int[1]
ci_upper_propane = t_test_propane$conf.int[2]
p_val_propane = t_test_propane$p.value
```
### 5.1 Confidence Interval
We estimate difference in gas prices (Northern Average - Southern Average) 
across the available weeks between 2021-2026 for Regular Unleaded Gasoline,
Diesel, .  
```{r eval = TRUE, echo = FALSE, message = T}
# Create a dataframe to store the results
summary_table = data.frame(
  "Fuel Type" = c(
    "Regular Unleaded Gasoline", 
    "Diesel", 
    "Auto Propane"
  ),
  "Mean Difference (cents/litre)" = c(
    round(mean_diff_gas, 2), 
    round(mean_diff_diesel, 2),
    round(mean_diff_propane, 2)
  ),
  "95% Confidence Interval" = c(
    paste0("(", round(ci_lower_gas, 2), ", ", round(ci_upper_gas, 2), ")"),
    paste0("(", round(ci_lower_diesel, 2), ", ", round(ci_upper_diesel, 2),
           ")"),
    paste0("(", round(ci_lower_propane, 2), ", ", round(ci_upper_propane, 2),
           ")")
  ),
  check.names = FALSE # This prevents R from putting periods in the column names
)

# Use kable to print a perfectly formatted table
kable(summary_table, 
      caption = "Average Price Differences: Northern vs. Southern Ontario",
      align = c("c", "c", "c"))
```
  
**Interpretation**:  
Based on data from the last five years, we are 95% confident that the true 
average difference in fuel prices between Northern and Southern Ontario falls 
within the following intervals (in cents per litre):

* **Regular Gas:** `r round(ci_lower_gas, 2)` to `r round(ci_upper_gas, 2)`
* **Diesel:** `r round(ci_lower_diesel, 2)` to `r round(ci_upper_diesel, 2)`
* **Auto Propane:** `r round(ci_lower_propane, 2)` to 
`r round(ci_upper_propane,2)`

### 5.2 Hypothesis Testing
To rigorously answer our research question regarding regional price differences,
we conducted a paired two-sample t-test for each of our three fuel types. We set
our significance level at $\alpha = 0.05$.

**Hypotheses:**
Let $\mu_d$ represent the true mean difference in weekly fuel prices between 
Northern and Southern Ontario ($\mu_{North} - \mu_{South}$).  

* **Null Hypothesis ($H_0$):** $\mu_d = 0$. There is no true mean difference in 
fuel prices between Northern and Southern Ontario. (Any observed difference is 
due to random market fluctuations).
* **Alternative Hypothesis ($H_1$):** $\mu_d \neq 0$. There is a true, 
statistically significant mean difference in fuel prices between Northern and 
Southern Ontario.

**Results:**

```{r eval = TRUE, echo = FALSE, message = FALSE}
# Create a dataframe for the Hypothesis Testing results
ht_table = data.frame(
  "Fuel Type" = c(
    "Regular Unleaded Gasoline", 
    "Diesel", 
    "Auto Propane"
  ),
  "p-value" = c(
    format.pval(p_val_gas, digits = 3, eps = 0.001),
    format.pval(p_val_diesel, digits = 3, eps = 0.001),
    format.pval(p_val_propane, digits = 3, eps = 0.001)
  ),
  check.names = FALSE
)

# Print the table using kable
kable(ht_table, 
      caption = "Paired T-Test Results: Regional Price Differences",
      align = c("l", "c", "c"))
```

**Conclusion:**   
As shown in the table above, the calculated p-values for all four fuel 
types—Regular Unleaded Gasoline, Diesel, Auto Propane, and Compressed Natural 
Gas—are strictly less than our significance level of 0.05. Because of this, 
**the price differences for every fuel category are statistically significant,** 
and **we confidently reject the null hypothesis ($H_0$) in all cases.**

## Section 6: Bootstrapping
While the paired t-tests in Section 5 provided strong evidence of a regional 
price divide, parametric tests rely on the assumption that the sample means are 
normally distributed. Real-world commodities like fuel are subject to sudden 
supply shocks and extreme price volatility, which can introduce outlines into 
the data. 

To ensure our findings are robust and not skewed by these potential outliers, we
utilize a non-parametric approach: **Bootstrapping**. By resampling our modern 
(2021-2026) paired regional price differences 1,000 times with replacement, we 
can generate an empirical distribution of the mean difference without relying on
assumptions of normality.

```{r eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
set.seed(57)  # Set a seed for exact reproducibility
B = 1000      # Number of bootstrap iterations

# Create a custom function using the tidyverse method taught in class
get_boot_ci = function(df) {
  
  # Define the single bootstrap step as taught in lecture
  boot_function = function() {
    # Resample the dataframe 
    boot_data = df %>% slice_sample(n = nrow(df), replace = TRUE)
    
    # Calculate the mean difference
    boot_mean = boot_data %>% summarize(cal.s = mean(Reg_Diff, na.rm = TRUE))
    
    # Extract the value
    return(boot_mean[1,1])
  }
  
  # Replicate the process B times
  boot_means = replicate(B, boot_function())
  
  # Extract the 2.5% and 97.5% percentiles for the 95% CI
  ci = quantile(boot_means, probs = c(0.025, 0.975))
  return(c(lower = ci[1], upper = ci[2]))
}

# Apply the function to all three fuel types using the dataframes we made in 
# Section 5
boot_gas = get_boot_ci(regional_compare_gas)
boot_diesel = get_boot_ci(regional_compare_diesel)
boot_propane = get_boot_ci(regional_compare_propane)
```

### 6.1 Bootstrap Confidence Intervals vs. Calssical Intervals
```{r eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
boot_table = data.frame(
  "Fuel Type" = c("Regular Unleaded Gasoline", "Diesel", "Auto Propane"),
  "Classical 95% CI (T-Test)" = c(
    paste0("(", round(ci_lower_gas, 2), ", ", round(ci_upper_gas, 2), ")"),
    paste0("(", round(ci_lower_diesel, 2), ", ", 
           round(ci_upper_diesel, 2), ")"),
    paste0("(", round(ci_lower_propane, 2), ", ", 
           round(ci_upper_propane, 2), ")")
  ),
  "Bootstrap 95% CI" = c(
    paste0("(", round(boot_gas["lower.2.5%"], 2), ", ",
           round(boot_gas["upper.97.5%"], 2), ")"),
    paste0("(", round(boot_diesel["lower.2.5%"], 2), ", ",
           round(boot_diesel["upper.97.5%"], 2), ")"),
    paste0("(", round(boot_propane["lower.2.5%"], 2), ", ",
           round(boot_propane["upper.97.5%"], 2), ")")
  ),
  check.names = FALSE
)

kable(boot_table, caption = 
        "Comparison of Classical and Bootstrapped 95% Confidence Intervals",
      align = c("l", "c", "c"))
```

### 6.2 Visualization of Bootstarpped Means
To better conceptualize this process, we can visualize the 1,000 resampled mean 
differences for Regular Unleaded Gasoline. The red dashed lines represent the 
boundaries of our 95% Bootstrap Confidence Interval.

```{r eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE, fig.align='center', fig.width=6, fig.height=3.5}
# 1. Create a function to extract the 1000 raw means for plotting
get_boot_means = function(df) {
  boot_func = function() {
    boot_data = df %>% slice_sample(n = nrow(df), replace = TRUE)
    boot_mean = boot_data %>% summarize(cal.s = mean(Reg_Diff, na.rm = TRUE))
    return(boot_mean[1,1])
  }
  return(replicate(B, boot_func()))
}

# 2. Generate the 1,000 means for the three fuel types
boot_means_gas = get_boot_means(regional_compare_gas)
boot_means_diesel = get_boot_means(regional_compare_diesel)
boot_means_propane = get_boot_means(regional_compare_propane)

# 3. Combine everything into one large dataframe AND set specific order (Levels)
plot_data = data.frame(
  Fuel_Type = factor(
    rep(c("Regular Unleaded Gasoline", "Diesel", "Auto Propane"), each = B),
    levels = c("Regular Unleaded Gasoline", "Diesel", "Auto Propane") 
    # This forces the left-to-right order
  ),
  Means = c(boot_means_gas, boot_means_diesel, boot_means_propane)
)

# 4. Create a dataframe for the Confidence Intervals with the same ordered 
#factor
ci_data = data.frame(
  Fuel_Type = factor(
    c("Regular Unleaded Gasoline", "Diesel", "Auto Propane"),
    levels = c("Regular Unleaded Gasoline", "Diesel", "Auto Propane")
  ),
  Lower = c(boot_gas[1], boot_diesel[1], boot_propane[1]),
  Upper = c(boot_gas[2], boot_diesel[2], boot_propane[2])
)

# 5. Plot using facet_wrap, adding colors and forcing 3 columns (left-to-right)
ggplot(plot_data, aes(x = Means, fill = Fuel_Type)) +
  geom_histogram(bins = 40, color = "black", linewidth = 0.2, alpha = 0.8,
                 show.legend = FALSE) +
  geom_vline(data = ci_data, aes(xintercept = Lower), color = "red",
             linetype = "dashed", linewidth = .4) +
  geom_vline(data = ci_data, aes(xintercept = Upper), color = "red",
             linetype = "dashed", linewidth = .4) +
  scale_fill_manual(values = c(
    "Regular Unleaded Gasoline" = "lightblue", 
    "Diesel" = "orange", 
    "Auto Propane" = "lightgreen"
  )) +
  facet_wrap(~ Fuel_Type, scales = "free_x", ncol = 3) + 
  labs(
    title = "Bootstrap Distributions of Mean Regional Differences",
    subtitle = "Northern Avg - Southern Avg (1,000 Resamples)",
    x = "Mean Difference (cents/litre)", 
    y = "Frequency"
  ) +
  theme_minimal(base_size = 14, base_family = "serif") +
  theme(strip.text = element_text(face = "bold", size = 10))
```

### 6.3 Interpretation
As demonstrated in the comparison table, our non-parametric Bootstrap confidence
intervals are nearly identical to the classical parametric intervals calculated 
in Section 5. Furthermore, the visualized distributions of the resampled means 
for all three fuel types exhibit clear, bell-shaped curves that sit entirely 
above zero.

## Section 7: Regression Analysis
In order to better understand the fuel price difference between the cities, we fit a regression model with time and city as predictors. From the weekly fuel price plot in Section 4, we can clearly see the correlation between time and fuel price is non-linear. Thus, we fit a polynomial regression model.  

```{r echo = FALSE}
# # Visualization on polynomial fitting
# plot_data = regular_gas_data %>%
#   filter(Date >= as.Date("2021-01-01")) %>%
#   mutate(Time_Index = as.numeric(Date - min(Date))) %>%
#   drop_na(Ontario_Average)
# 
# ggplot(plot_data, aes(x = Time_Index, y = Ontario_Average)) + 
#   geom_point() + 
#   geom_smooth(method = "lm", formula = y ~ poly(x, 1), 
#               aes(color = "Deg 1"), se = FALSE, linewidth = 1) + 
#   geom_smooth(method = "lm", formula = y ~ poly(x, 2), 
#               aes(color = "Deg 2"), se = FALSE, linewidth = 1) + 
#   geom_smooth(method = "lm", formula = y ~ poly(x, 3), 
#               aes(color = "Deg 3"), se = FALSE, linewidth = 1) +
#   geom_smooth(method = "lm", formula = y ~ poly(x, 5), 
#               aes(color = "Deg 5"), se = FALSE, linewidth = 1) +
#   labs(
#     title = "Testing Polynomial Fits on Fuel Prices",
#     x = "Week",
#     y = "Ontario Average Fuel Price (cents/litre)",
#     color = "Model Fit"
#   ) +
#   theme_minimal()
```

  

```{r echo = FALSE}
############# Put This in Data cleaning section!
regular_gas_data = regular_gas_data %>%
  filter(Date >= as.Date("2021-01-01")) %>%
  select(Date, Ottawa, Windsor, London, Peterborough, Sudbury, North_Bay, Parry_Sound, Sault_Saint_Marie, St_Catharines, Thunder_Bay, Toronto_East, Toronto_West, Timmins, Kenora) %>%
  pivot_longer(cols = -Date, names_to = "City", values_to = "Price") %>%
  drop_na() %>%
  mutate(Time_Index = as.numeric(Date - min(Date)))

diesel_data = diesel_data %>%
  filter(Date >= as.Date("2021-01-01")) %>%
  select(Date, Ottawa, Windsor, London, Peterborough, Sudbury, North_Bay, Parry_Sound, 
         Sault_Saint_Marie, St_Catharines, Thunder_Bay, Toronto_East, Toronto_West, 
         Timmins, Kenora) %>%
  pivot_longer(cols = -Date, names_to = "City", values_to = "Price") %>%
  drop_na() %>%
  mutate(Time_Index = as.numeric(Date - min(Date)))

auto_propane_data = auto_propane_data %>%
  filter(Date >= as.Date("2021-01-01")) %>%
  select(Date, Ottawa, Windsor, London, Thunder_Bay, Toronto_East, Toronto_West, 
         Kenora) %>%
  pivot_longer(cols = -Date, names_to = "City", values_to = "Price") %>%
  drop_na() %>%
  mutate(Time_Index = as.numeric(Date - min(Date)))

```



To prevent overfitting, we build the full model of maximum degree 5, and use cross-validation to determine the best degree model. To make sure the model perform well throughout this time period, we use 5-fold cross-validation.

```{r echo=FALSE}

# Data split
regular = regular_gas_data %>% 
  mutate(group_ind = sample(c(1:5), size = nrow(regular_gas_data), replace = TRUE))
diesel = diesel_data %>% 
  mutate(group_ind = sample(c(1:5), size = nrow(regular_gas_data), replace = TRUE))
propane = auto_propane_data %>% 
  mutate(group_ind = sample(c(1:5), size = nrow(auto_propane_data), replace = TRUE))


regular_mse = vector()
diesel_mse = vector()
propane_mse = vector()

for(d in 1:5) {
  mse_fold = vector() 

  for(i in 1:5) {
    
    train_set = regular %>% filter(group_ind != i)
    validation_set = regular %>% filter(group_ind == i)
    
    m_train = lm(Price ~ City + poly(Time_Index, d), data = train_set)
    y_hat = predict(m_train, newdata = validation_set)
    mse_fold[i] = mean((validation_set$Price - y_hat)^2, na.rm = TRUE)
  }
  
  # Take average as final score
  regular_mse[d] = mean(mse_fold)
}


cv_result_regular = data.frame(
  "Polynomial Degree" = c("Degree 1", "Degree 2", "Degree 3", "Degree 4", "Degree 5"),
  "Cross-Validated MSE" = regular_mse,
  check.names = FALSE
)

for(d in 1:5) {
  mse_fold = vector() 

  for(i in 1:5) {
    
    train_set = diesel %>% filter(group_ind != i)
    validation_set = diesel %>% filter(group_ind == i)
    
    m_train = lm(Price ~ City + poly(Time_Index, d), data = train_set)
    y_hat = predict(m_train, newdata = validation_set)
    mse_fold[i] = mean((validation_set$Price - y_hat)^2, na.rm = TRUE)
  }
  
  diesel_mse[d] = mean(mse_fold)
}

cv_result_diesel = data.frame(
  "Polynomial Degree" = c("Degree 1", "Degree 2", "Degree 3", "Degree 4", "Degree 5"),
  "Cross-Validated MSE" = diesel_mse,
  check.names = FALSE
)

for(d in 1:5) {
  mse_fold = vector() 

  for(i in 1:5) {
    # Explicitly defining Train and Validation sets per fold
    train_set = propane %>% filter(group_ind != i)
    validation_set = propane %>% filter(group_ind == i)
    
    m_train = lm(Price ~ City + poly(Time_Index, d), data = train_set)
    y_hat = predict(m_train, newdata = validation_set)
    mse_fold[i] = mean((validation_set$Price - y_hat)^2, na.rm = TRUE)
  }
  
  propane_mse[d] = mean(mse_fold)
}

cv_result_propane = data.frame(
  "Polynomial Degree" = c("Degree 1", "Degree 2", "Degree 3", "Degree 4", "Degree 5"),
  "Cross-Validated MSE" = propane_mse,
  check.names = FALSE
)



kable(cv_result_regular, caption = "Cross-Validation Error (Regular Gas)")
kable(cv_result_diesel, caption = "Cross-Validation Error (Diesel)")
kable(cv_result_propane, caption = "Cross-Validation Error (Auto Propane)")

plot_cv_data = data.frame(
  Degree = rep(1:5, 3),
  MSE = c(regular_mse, diesel_mse, propane_mse),
  Fuel_Type = rep(c("Regular Gas", "Diesel", "Auto Propane"), each = 5)
)

ggplot(plot_cv_data, aes(x = Degree, y = MSE, color = Fuel_Type, group = Fuel_Type)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  scale_color_manual(values = c("Regular Gas" = "blue", 
                                "Diesel" = "red", 
                                "Auto Propane" = "orange")) +
  labs(
    title = "Cross-Validation Error by Polynomial Degree",
    x = "Polynomial Degree",
    y = "MSE",
    color = "Fuel Type"
  )
```
  
We can see for all fuel types, from degree 1 to degree 3, MSE dropped significantly because of the increase in model complexity. However, after degree 3, the extra accuracy is marginal. Since we prefer simpler model at similar accuracy, we choose degree 3 model as the final model.

```{r echo=FALSE}
# Final models
regular_fit = lm(Price ~ City + poly(Time_Index, 3), data = regular_gas_data)
diesel_fit = lm(Price ~ City + poly(Time_Index, 3), data = diesel_data)
propane_fit = lm(Price ~ City + poly(Time_Index, 3), data = auto_propane_data)
summary(regular_fit)
summary(diesel_fit)
summary(propane_fit)
```



## Section 8: Conclusion

\pagebreak

## Section 9: Appendix

### Section 1: Introduction

### Section 2: Data Description

### Section 3: Data Preparation/Cleaning

```{r eval = FALSE}
# Import needed libraries
library(tidyverse)
library(dplyr)
library(lubridate)
library(knitr)
library(ggplot2)
```

```{r eval = FALSE}
# Get the data
raw_data = read.csv("data/fueltypesall.csv")
```

```{r eval = FALSE}
# Rename the columns to full English
d = raw_data %>%
  rename(
    Toronto_West = Toronto.West.Ouest,
    Toronto_East = Toronto.East.Est,
    St_Catharines = St..Catharine.s,
    Sault_Saint_Marie = Sault.Saint.Marie,
    Thunder_Bay = Thunder.Bay,
    North_Bay = North.Bay,
    Parry_Sound = Parry.Sound,
    Ontario_Average = Ontario.Average.Moyenne.provinciale,
    Southern_Ontario_Average = Southern.Average.Moyenne.du.sud.de.l.Ontario,
    Northern_Ontario_Average = Northern.Average.Moyenne.du.nord.de.l.Ontario,
    Fuel_Type = Fuel.Type
  )
```

```{r eval = FALSE}
# Show the column name changes in a comparison table
old_cols = names(raw_data)
new_cols = names(d)

names_comparison = data.frame(
  Before = old_cols,
  After = new_cols,
  stringsAsFactors = FALSE 
)
kable(names_comparison, caption = "Old Column Names vs. New Column Names")
```

```{r eval = FALSE}
# Drop the column Type.de.carburant
d = d %>% 
  select(!Type.de.carburant)
```

```{r eval = FALSE}
# Convert 0 values to NA
d = d %>%
  mutate(across(where(is.numeric), ~ na_if(., 0)))
```

```{r eval = FALSE}
# Change the Date column type from char to date
d = d %>%
  mutate(Date = as.Date(Date))
```

```{r eval = FALSE}
# Separate the data set into the six fuel categories and drop the Fuel_Type 
# column for each new data set
regular_gas_data = d %>%
  filter(Fuel_Type == "Regular Unleaded Gasoline") %>%
  select(!Fuel_Type)

diesel_data = d %>%
  filter(Fuel_Type == "Diesel") %>%
  select(!Fuel_Type)

natural_gas_data = d %>%
  filter(Fuel_Type == "Compressed Natural Gas") %>%
  select(!Fuel_Type)

mid_gas_data = d %>%
  filter(Fuel_Type == "Mid-Grade Gasoline") %>%
  select(!Fuel_Type) 

premium_gas_data = d %>%
  filter(Fuel_Type == "Premium Gasoline") %>%
  select(!Fuel_Type)

auto_propane_data = d %>%
  filter(Fuel_Type == "Auto Propane") %>%
  select(!Fuel_Type) 
```

### Section 4: Exploratory Data Analysis
```{r eval = FALSE}
d %>% 
  ggplot(aes(x = Date, y = Ontario_Average, color = Fuel_Type)) + 
  geom_line() +
  labs(
    title = "Weekly Fuel Prices in Ontario by Fuel Type",
    x = "Week",
    y = "Average Fuel Price (cents/litre)",
    color = "Fuel Type")

```

### Section 5: Statistical Inference
```{r eval = F, echo = T, message = FALSE}
### Filters the dataset for recent years (2021+) and runs paired t-tests to 
### calculate the average regional price difference, confidence intervals, and 
### statistical significance for gasoline, diesel, and auto propane.

# REGULAR GASOLINE
regional_compare_gas = regular_gas_data %>%
  filter(Date >= as.Date("2021-01-01")) %>%
  select(Date, Northern_Ontario_Average, Southern_Ontario_Average) %>%
  drop_na() %>%
  mutate(Reg_Diff = Northern_Ontario_Average - Southern_Ontario_Average)

t_test_gas = t.test(regional_compare_gas$Northern_Ontario_Average, 
                    regional_compare_gas$Southern_Ontario_Average,
                    paired = TRUE)
mean_diff_gas = t_test_gas$estimate
ci_lower_gas = t_test_gas$conf.int[1]
ci_upper_gas = t_test_gas$conf.int[2]
p_val_gas = t_test_gas$p.value

# DIESEL
regional_compare_diesel = diesel_data %>%
  filter(Date >= as.Date("2021-01-01")) %>%
  select(Date, Northern_Ontario_Average, Southern_Ontario_Average) %>%
  drop_na() %>%
  mutate(Reg_Diff = Northern_Ontario_Average - Southern_Ontario_Average)

t_test_diesel = t.test(regional_compare_diesel$Northern_Ontario_Average,
                       regional_compare_diesel$Southern_Ontario_Average,
                       paired = TRUE)
mean_diff_diesel = t_test_diesel$estimate
ci_lower_diesel = t_test_diesel$conf.int[1]
ci_upper_diesel = t_test_diesel$conf.int[2]
p_val_diesel = t_test_diesel$p.value

# AUTO PROPANE
regional_compare_propane = auto_propane_data %>%
  filter(Date >= as.Date("2021-01-01")) %>%
  select(Date, Northern_Ontario_Average, Southern_Ontario_Average) %>%
  drop_na() %>%
  mutate(Reg_Diff = Northern_Ontario_Average - Southern_Ontario_Average)

t_test_propane = t.test(regional_compare_propane$Northern_Ontario_Average,
                        regional_compare_propane$Southern_Ontario_Average,
                        paired = TRUE)
mean_diff_propane = t_test_propane$estimate
ci_lower_propane = t_test_propane$conf.int[1]
ci_upper_propane = t_test_propane$conf.int[2]
p_val_propane = t_test_propane$p.value
```

```{r eval = F, echo = T, message = F}
### Compiles the calculated mean differences and 95% confidence intervals into a
### dataframe and generates a cleanly formatted summary table using the kable 
## function.

# Create a dataframe to store the results
summary_table = data.frame(
  "Fuel Type" = c(
    "Regular Unleaded Gasoline", 
    "Diesel", 
    "Auto Propane"
  ),
  "Mean Difference (cents/litre)" = c(
    round(mean_diff_gas, 2), 
    round(mean_diff_diesel, 2),
    round(mean_diff_propane, 2)
  ),
  "95% Confidence Interval" = c(
    paste0("(", round(ci_lower_gas, 2), ", ", round(ci_upper_gas, 2), ")"),
    paste0("(", round(ci_lower_diesel, 2), ", ", round(ci_upper_diesel, 2),
           ")"),
    paste0("(", round(ci_lower_propane, 2), ", ", round(ci_upper_propane, 2),
           ")")
  ),
  check.names = FALSE # This prevents R from putting periods in the column names
)

# Use kable to print a perfectly formatted table
kable(summary_table, 
      caption = "Average Price Differences: Northern vs. Southern Ontario",
      align = c("c", "c", "c"))
```

```{r eval = F, echo = T, message = FALSE}
### Compiles the p-values from the paired t-tests into a summary table to 
### evaluate the statistical significance of regional price differences across 
### the three fuel types.

# Dataframe for the Hypothesis Testing results
ht_table = data.frame(
  "Fuel Type" = c(
    "Regular Unleaded Gasoline", 
    "Diesel", 
    "Auto Propane"
  ),
  "p-value" = c(
    format.pval(p_val_gas, digits = 3, eps = 0.001),
    format.pval(p_val_diesel, digits = 3, eps = 0.001),
    format.pval(p_val_propane, digits = 3, eps = 0.001)
  ),
  check.names = FALSE
)

# Table
kable(ht_table, 
      caption = "Paired T-Test Results: Regional Price Differences",
      align = c("l", "c", "c"))
```

### Section 6: Bootstrapping
```{r eval = F, echo = T, message = FALSE, warning = FALSE}
### Mimics the concept of repeated sampling by using the bootstrap method to 
### generate distributions and 95% confidence intervals for regional 
### price differences across the three fuel types.

set.seed(57)
B = 1000      # Number of bootstrap iterations

get_boot_ci = function(df) {
  
  boot_function = function() {
    boot_data = df %>% slice_sample(n = nrow(df), replace = TRUE)
    boot_mean = boot_data %>% summarize(cal.s = mean(Reg_Diff, na.rm = TRUE))
    return(boot_mean[1,1])
  }
  
  # Replicate the process B times
  boot_means = replicate(B, boot_function())
  
  # 2.5% and 97.5% percentiles for the 95% CI
  ci = quantile(boot_means, probs = c(0.025, 0.975))
  return(c(lower = ci[1], upper = ci[2]))
}

# Apply the function to all three fuel types using the dataframes we made in 
# Section 5
boot_gas = get_boot_ci(regional_compare_gas)
boot_diesel = get_boot_ci(regional_compare_diesel)
boot_propane = get_boot_ci(regional_compare_propane)
```

```{r eval = F, echo = T, message = FALSE, warning = FALSE}
### Compiles the 95% confidence intervals from both the classical t-tests and 
### the bootstrap resampling into a comparative dataframe to demonstrate the 
### stability of the regional price difference estimates across all three fuel 
### types.

boot_table = data.frame(
  "Fuel Type" = c("Regular Unleaded Gasoline", "Diesel", "Auto Propane"),
  "Classical 95% CI (T-Test)" = c(
    paste0("(", round(ci_lower_gas, 2), ", ", round(ci_upper_gas, 2), ")"),
    paste0("(", round(ci_lower_diesel, 2), ", ", 
           round(ci_upper_diesel, 2), ")"),
    paste0("(", round(ci_lower_propane, 2), ", ", 
           round(ci_upper_propane, 2), ")")
  ),
  "Bootstrap 95% CI" = c(
    paste0("(", round(boot_gas["lower.2.5%"], 2), ", ",
           round(boot_gas["upper.97.5%"], 2), ")"),
    paste0("(", round(boot_diesel["lower.2.5%"], 2), ", ",
           round(boot_diesel["upper.97.5%"], 2), ")"),
    paste0("(", round(boot_propane["lower.2.5%"], 2), ", ",
           round(boot_propane["upper.97.5%"], 2), ")")
  ),
  check.names = FALSE
)

kable(boot_table, caption = 
        "Comparison of Classical and Bootstrapped 95% Confidence Intervals",
      align = c("l", "c", "c"))
```

```{r eval = F, echo = T, message = FALSE, warning = FALSE, fig.align='center', fig.width=6, fig.height=3.5}
#Creates a function to extract the 1000 raw means for plotting
get_boot_means = function(df) {
  boot_func = function() {
    boot_data = df %>% slice_sample(n = nrow(df), replace = TRUE)
    boot_mean = boot_data %>% summarize(cal.s = mean(Reg_Diff, na.rm = TRUE))
    return(boot_mean[1,1])
  }
  return(replicate(B, boot_func()))
}

# Generates the 1,000 means for the three fuel types
boot_means_gas = get_boot_means(regional_compare_gas)
boot_means_diesel = get_boot_means(regional_compare_diesel)
boot_means_propane = get_boot_means(regional_compare_propane)

#Combines everything into one large dataframe AND set specific order (Levels)
plot_data = data.frame(
  Fuel_Type = factor(
    rep(c("Regular Unleaded Gasoline", "Diesel", "Auto Propane"), each = B),
    levels = c("Regular Unleaded Gasoline", "Diesel", "Auto Propane") 
  ),
  Means = c(boot_means_gas, boot_means_diesel, boot_means_propane)
)

#Creates a dataframe for the Confidence Intervals with the same ordered factor
ci_data = data.frame(
  Fuel_Type = factor(
    c("Regular Unleaded Gasoline", "Diesel", "Auto Propane"),
    levels = c("Regular Unleaded Gasoline", "Diesel", "Auto Propane")
  ),
  Lower = c(boot_gas[1], boot_diesel[1], boot_propane[1]),
  Upper = c(boot_gas[2], boot_diesel[2], boot_propane[2])
)

#Plot using facet_wrap, adding colors and forcing 3 columns (left-to-right)
ggplot(plot_data, aes(x = Means, fill = Fuel_Type)) +
  geom_histogram(bins = 40, color = "black", linewidth = 0.2, alpha = 0.8,
                 show.legend = FALSE) +
  geom_vline(data = ci_data, aes(xintercept = Lower), color = "red",
             linetype = "dashed", linewidth = .4) +
  geom_vline(data = ci_data, aes(xintercept = Upper), color = "red",
             linetype = "dashed", linewidth = .4) +
  scale_fill_manual(values = c(
    "Regular Unleaded Gasoline" = "steelblue", 
    "Diesel" = "darkorange", 
    "Auto Propane" = "forestgreen"
  )) +
  facet_wrap(~ Fuel_Type, scales = "free_x", ncol = 3) + 
  labs(
    title = "Bootstrap Distributions of Mean Regional Differences",
    subtitle = "Northern Avg - Southern Avg (1,000 Resamples)",
    x = "Mean Difference (cents/litre)", 
    y = "Frequency"
  ) +
  theme_minimal(base_size = 14, base_family = "serif") +
  theme(strip.text = element_text(face = "bold", size = 10))
```

### Section 7: Regression Analysis
```{r eval=FALSE}

# Data split
regular = regular_gas_data %>% 
  mutate(group_ind = sample(c(1:5), size = nrow(regular_gas_data), replace = TRUE))
diesel = diesel_data %>% 
  mutate(group_ind = sample(c(1:5), size = nrow(regular_gas_data), replace = TRUE))
propane = auto_propane_data %>% 
  mutate(group_ind = sample(c(1:5), size = nrow(auto_propane_data), replace = TRUE))


regular_mse = vector()
diesel_mse = vector()
propane_mse = vector()

for(d in 1:5) {
  mse_fold = vector() 

  for(i in 1:5) {
    
    train_set = regular %>% filter(group_ind != i)
    validation_set = regular %>% filter(group_ind == i)
    
    m_train = lm(Price ~ City + poly(Time_Index, d), data = train_set)
    y_hat = predict(m_train, newdata = validation_set)
    mse_fold[i] = mean((validation_set$Price - y_hat)^2, na.rm = TRUE)
  }
  
  # Take average as final score
  regular_mse[d] = mean(mse_fold)
}


cv_result_regular = data.frame(
  "Polynomial Degree" = c("Degree 1", "Degree 2", "Degree 3", "Degree 4", "Degree 5"),
  "Cross-Validated MSE" = regular_mse,
  check.names = FALSE
)

for(d in 1:5) {
  mse_fold = vector() 

  for(i in 1:5) {
    
    train_set = diesel %>% filter(group_ind != i)
    validation_set = diesel %>% filter(group_ind == i)
    
    m_train = lm(Price ~ City + poly(Time_Index, d), data = train_set)
    y_hat = predict(m_train, newdata = validation_set)
    mse_fold[i] = mean((validation_set$Price - y_hat)^2, na.rm = TRUE)
  }
  
  diesel_mse[d] = mean(mse_fold)
}

cv_result_diesel = data.frame(
  "Polynomial Degree" = c("Degree 1", "Degree 2", "Degree 3", "Degree 4", "Degree 5"),
  "Cross-Validated MSE" = diesel_mse,
  check.names = FALSE
)

for(d in 1:5) {
  mse_fold = vector() 

  for(i in 1:5) {
    # Explicitly defining Train and Validation sets per fold
    train_set = propane %>% filter(group_ind != i)
    validation_set = propane %>% filter(group_ind == i)
    
    m_train = lm(Price ~ City + poly(Time_Index, d), data = train_set)
    y_hat = predict(m_train, newdata = validation_set)
    mse_fold[i] = mean((validation_set$Price - y_hat)^2, na.rm = TRUE)
  }
  
  propane_mse[d] = mean(mse_fold)
}

cv_result_propane = data.frame(
  "Polynomial Degree" = c("Degree 1", "Degree 2", "Degree 3", "Degree 4", "Degree 5"),
  "Cross-Validated MSE" = propane_mse,
  check.names = FALSE
)



kable(cv_result_regular, caption = "Cross-Validation Error (Regular Gas)")
kable(cv_result_diesel, caption = "Cross-Validation Error (Diesel)")
kable(cv_result_propane, caption = "Cross-Validation Error (Auto Propane)")

plot_cv_data = data.frame(
  Degree = rep(1:5, 3),
  MSE = c(regular_mse, diesel_mse, propane_mse),
  Fuel_Type = rep(c("Regular Gas", "Diesel", "Auto Propane"), each = 5)
)

ggplot(plot_cv_data, aes(x = Degree, y = MSE, color = Fuel_Type, group = Fuel_Type)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  scale_color_manual(values = c("Regular Gas" = "blue", 
                                "Diesel" = "red", 
                                "Auto Propane" = "orange")) +
  labs(
    title = "Cross-Validation Error by Polynomial Degree",
    x = "Polynomial Degree",
    y = "MSE",
    color = "Fuel Type"
  )
```


```{r eval=FALSE}
# Final models
regular_fit = lm(Price ~ City + poly(Time_Index, 3), data = regular_gas_data)
diesel_fit = lm(Price ~ City + poly(Time_Index, 3), data = diesel_data)
propane_fit = lm(Price ~ City + poly(Time_Index, 3), data = auto_propane_data)
summary(regular_fit)
summary(diesel_fit)
summary(propane_fit)
```





### Section 8: Conclusion
