---
title: "Ontario Fuel Price Trends: 1990 to 2026"
author: "Michael Zhou, Nigel Li, Stella Pham"
date: "2026-03-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r eval = TRUE, echo = FALSE, message = FALSE}
library(tidyverse)
library(dplyr)
library(lubridate)
library(knitr)
library(ggplot2)
```

```{r eval = TRUE, echo = FALSE, message = FALSE}
raw_data = read.csv("data/fueltypesall.csv")
```

## Section 1: Introduction
### Background of the Data
This data set contains weekly retail prices for six fuel types: Regular Unleaded 
Gasoline, Mid‑Grade Gasoline, Premium Gasoline, Diesel, Auto Propane, and 
Compressed Natural Gas across 10 markets in Ontario. Prices are expressed in 
Canadian cents per litre, with compressed natural gas measured in cents per 
gasoline-equivalent litre. The data set spans from January 3, 1990 to March 16, 
2026 and is updated weekly with the latest retail fuel price data published by 
the Government of Ontario.

### Research Question
How does the price differ between cities in Ontario?

## Section 2: Data Description
The data set contains 11,252 weeks of data from 1990 to March 16, 2026, covering 
six fuel types. Each row represents a weekly average of fuel prices, so daily 
fluctuations are not captured. It also includes three regional average columns 
(Ontario, Northern Ontario, Southern Ontario), which do not correspond to the 
simple average of the included cities, indicating they were sourced externally.

## Section 3: Data Preparation/Cleaning
Prior to analysis, it is important to highlight certain features of the raw 
data set that affect how it should be handled:

### 1. Bilingual Structure of the Data Set
The data set contains both French and English labels, sometimes separated by 
slashes or presented in different columns. To maintain consistency, 
variables are standardized into a single language format before analysis.

We first rename the columns we keep.

```{r eval = TRUE, echo = FALSE, message = FALSE}
d = raw_data %>%
  rename(
    Toronto_West = Toronto.West.Ouest,
    Toronto_East = Toronto.East.Est,
    St_Catharines = St..Catharine.s,
    Sault_Saint_Marie = Sault.Saint.Marie,
    Thunder_Bay = Thunder.Bay,
    North_Bay = North.Bay,
    Parry_Sound = Parry.Sound,
    Ontario_Average = Ontario.Average.Moyenne.provinciale,
    Southern_Ontario_Average = Southern.Average.Moyenne.du.sud.de.l.Ontario,
    Northern_Ontario_Average = Northern.Average.Moyenne.du.nord.de.l.Ontario,
    Fuel_Type = Fuel.Type
  )
```

```{r eval = TRUE, echo = FALSE, message = TRUE}
old_cols = names(raw_data)
new_cols = names(d)

names_comparison = data.frame(
  Before = old_cols,
  After = new_cols,
  stringsAsFactors = FALSE 
)

library(knitr)
kable(names_comparison, caption = "Old Column Names vs. New Column Names")
```

```{r eval = TRUE, echo = FALSE, message = FALSE}
d = d %>% 
  select(!Type.de.carburant)
```

We also remove the column "Type.de.carburant" was removed because it was the 
same as the information in the Fuel_Type column. More specifically, it contained 
the fuel type names in French.

### 2. Handling of Zero Values
Values recorded as 0 in the city-level columns are interpreted as missing data, 
since these 0s do not represent actual prices. To prevent these zeros from 
affecting computations, all 0 values are converted to NA prior to analysis.

```{r eval = TRUE, echo = FALSE, message = FALSE}
d = d %>%
  mutate(across(where(is.numeric), ~ na_if(., 0)))
```

### 3. Conversion of the Date Column
The Week column, originally recorded as a character string, was converted to a 
Date type to ensure proper chronological ordering and accurate calculations. 
This allows for correct filtering, aggregation, and plotting over time, and 
ensures that weekly comparisons between cities are reliable.

```{r eval = TRUE, echo = FALSE, message = FALSE}
d = d %>%
  mutate(Date = as.Date(Date))
```

### 4. Categorization of the Data Set
As mentioned above, the data set includes six fuel types. To enable clearer 
and more accurate comparisons between cities, the data will be analyzed 
separately for each fuel type.

```{r eval = TRUE, echo = FALSE, message = FALSE}
regular_gas_data = d %>%
  filter(Fuel_Type == "Regular Unleaded Gasoline") %>%
  select(!Fuel_Type)

diesel_data = d %>%
  filter(Fuel_Type == "Diesel") %>%
  select(!Fuel_Type)

natural_gas_data = d %>%
  filter(Fuel_Type == "Compressed Natural Gas") %>%
  select(!Fuel_Type)

mid_gas_data = d %>%
  filter(Fuel_Type == "Mid-Grade Gasoline") %>%
  select(!Fuel_Type) 

premium_gas_data = d %>%
  filter(Fuel_Type == "Premium Gasoline") %>%
  select(!Fuel_Type)

auto_propane_data = d %>%
  filter(Fuel_Type == "Auto Propane") %>%
  select(!Fuel_Type) 
```

## Section 4: Exploratory Data Analysis
### Visualizations

We want

```{r eval = TRUE, echo = FALSE, warning = FALSE}
d %>% 
  ggplot(aes(x = Date, y = Ontario_Average, color = Fuel_Type)) + 
  geom_line() +
  labs(
    title = "Weekly Fuel Prices in Ontario by Fuel Type",
    x = "Week",
    y = "Average Fuel Price (cents/litre)",
    color = "Fuel Type")
```

## Section 5: Statistical Inference
### Confidence Interval

### Hypothesis Testing

## Section 6: Bootstrapping

## Section 7: Regression Analysis

## Section 8: Model Validation

## Section 9: Conclusion

\pagebreak

## Section 10: Appendix

### Section 1: Introduction

### Section 2: Data Description

### Section 3: Data Preparation/Cleaning

```{r eval = FALSE}
# Import needed libraries
library(tidyverse)
library(dplyr)
library(lubridate)
library(knitr)
library(ggplot2)
```

```{r eval = FALSE}
# Get the data
raw_data = read.csv("data/fueltypesall.csv")
```

```{r eval = FALSE}
# Rename the columns to full English
d = raw_data %>%
  rename(
    Toronto_West = Toronto.West.Ouest,
    Toronto_East = Toronto.East.Est,
    St_Catharines = St..Catharine.s,
    Sault_Saint_Marie = Sault.Saint.Marie,
    Thunder_Bay = Thunder.Bay,
    North_Bay = North.Bay,
    Parry_Sound = Parry.Sound,
    Ontario_Average = Ontario.Average.Moyenne.provinciale,
    Southern_Ontario_Average = Southern.Average.Moyenne.du.sud.de.l.Ontario,
    Northern_Ontario_Average = Northern.Average.Moyenne.du.nord.de.l.Ontario,
    Fuel_Type = Fuel.Type
  )
```

```{r eval = FALSE}
# Show the column name changes in a comparison table
old_cols = names(raw_data)
new_cols = names(d)

names_comparison = data.frame(
  Before = old_cols,
  After = new_cols,
  stringsAsFactors = FALSE 
)
kable(names_comparison, caption = "Old Column Names vs. New Column Names")
```

```{r eval = FALSE}
# Drop the column Type.de.carburant
d = d %>% 
  select(!Type.de.carburant)
```

```{r eval = FALSE}
# Convert 0 values to NA
d = d %>%
  mutate(across(where(is.numeric), ~ na_if(., 0)))
```

```{r eval = FALSE}
# Change the Date column type from char to date
d = d %>%
  mutate(Date = as.Date(Date))
```

```{r eval = FALSE}
# Separate the data set into the six fuel categories and drop the Fuel_Type column for each new data set
regular_gas_data = d %>%
  filter(Fuel_Type == "Regular Unleaded Gasoline") %>%
  select(!Fuel_Type)

diesel_data = d %>%
  filter(Fuel_Type == "Diesel") %>%
  select(!Fuel_Type)

natural_gas_data = d %>%
  filter(Fuel_Type == "Compressed Natural Gas") %>%
  select(!Fuel_Type)

mid_gas_data = d %>%
  filter(Fuel_Type == "Mid-Grade Gasoline") %>%
  select(!Fuel_Type) 

premium_gas_data = d %>%
  filter(Fuel_Type == "Premium Gasoline") %>%
  select(!Fuel_Type)

auto_propane_data = d %>%
  filter(Fuel_Type == "Auto Propane") %>%
  select(!Fuel_Type) 
```

### Section 4: Exploratory Data Analysis
```{r eval = FALSE}
d %>% 
  ggplot(aes(x = Date, y = Ontario_Average, color = Fuel_Type)) + 
  geom_line() +
  labs(
    title = "Weekly Fuel Prices in Ontario by Fuel Type",
    x = "Week",
    y = "Average Fuel Price (cents/litre)",
    color = "Fuel Type")

```


### Section 5: Statistical Inference

### Section 6: Bootstrapping

### Section 7: Regression Analysis

### Section 8: Model Validation

### Section 9: Conclusion
